---
title: "COVID-19"
author: "Christof Seiler"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    theme: spacelab
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height=4, fig.width=8)
library("tidyverse")
library("magrittr")
library("lubridate")
library("ggthemes")
```

# Data

Load GitHub data package.

```{r warning=FALSE, message=FALSE}
devtools::install_github("RamiKrispin/coronavirus")
library("coronavirus")
```

This repository is updated daily. The data frame is in variable `coronavirus`.

```{r}
data("coronavirus")
head(coronavirus)
tail(coronavirus)
```

Summary.

```{r}
coronavirus %>% 
  group_by(Country.Region, type) %>%
  summarize(total_cases = sum(cases)) %>%
  arrange(-total_cases)
```

Remove zero counts.

```{r}
coronavirus %<>% filter(cases > 0)
```

# Visualize Development

Here are the daily development of the following 5 nations:

* China in the province Hubei
* South Korea
* Italy
* Switzerland
* The Netherlands

Also some comparisons between the Netherlands and other countries.

## Hubei Province in China

```{r}
plot_daily = function(df) {
  ggplot(df, aes(date, cases, color = type)) + 
    geom_point() +
    geom_line() + 
    facet_wrap(~Province.State) + 
    scale_x_date(date_minor_breaks = "1 day", 
                 date_breaks = "1 week", 
                 date_labels = "%b %d") + 
    scale_color_few()
}
df_china = coronavirus %>% filter(
  Country.Region == "China",
  Province.State == "Hubei"
  )
plot_daily(df_china) + ggtitle("Mainland China")
```

## South Korea

```{r}
df_south_korea = coronavirus %>% filter(Country.Region == "Korea, South")
plot_daily(df_south_korea) + ggtitle("South Korea")
```

## Italy

```{r}
df_italy = coronavirus %>% filter(Country.Region == "Italy")
plot_daily(df_italy) + ggtitle("Italy")
```

## Switzerland

```{r}
df_switzerland = coronavirus %>% filter(Country.Region == "Switzerland")
plot_daily(df_switzerland) + 
  ggtitle("Switzerland") +
   scale_x_date(date_minor_breaks = "1 day", 
                 date_breaks = "7 day", 
                 date_labels = "%b %d")
```

## The Netherlands

```{r}
df_switzerland = coronavirus %>% filter(Country.Region == "Netherlands")
plot_daily(df_switzerland) + 
  ggtitle("Netherlands") +
   scale_x_date(date_minor_breaks = "1 day", 
                 date_breaks = "7 day", 
                 date_labels = "%b %d")
```

## Compare to Neighbors

Compare daily developments to our neighboring countries.

```{r}
plot_compare = function(df) {
  ggplot(df, aes(date, cases, color = Country.Region)) + 
    geom_point() +
    geom_line() + 
    scale_x_date(date_minor_breaks = "1 day", 
                 date_breaks = "1 week", 
                 date_labels = "%b %d") +
    scale_color_few()
}
neighbors = c("Netherlands", "Belgium", "Germany", "Luxembourg")
df_neighbors = coronavirus %>% filter(
  Country.Region %in% neighbors & type == "confirmed"
  )
plot_compare(df_neighbors)
```

## Compare to Italy

Compare the Netherlands to Italy.

```{r}
neighbors = c("Netherlands", "Italy")
df_neighbors = coronavirus %>% filter(
  Country.Region %in% neighbors & type == "confirmed"
  )
plot_compare(df_neighbors)
```

## Compare to Switzerland

Compare the Netherlands to Switzerland.

```{r}
neighbors = c("Netherlands", "Switzerland")
df_neighbors = coronavirus %>% filter(
  Country.Region %in% neighbors & type == "confirmed"
  )
plot_compare(df_neighbors) +
  scale_x_date(date_minor_breaks = "1 day", 
                 date_breaks = "7 day", 
                 date_labels = "%b %d")
```

# Visualize Align Countries

Align time series with respect to days from first confirmed case in respective country. 

* First, find patient zero for each country.
* Second, plot aligned number of cases to compare the relative progress between country.

## Patient Zero

Patient zero for each country.

```{r}
add_patient_n = function(coronavirus, n_cases = 1) {
  df = coronavirus
  df_patient_n = df %>% 
    group_by(Country.Region) %>%
    arrange(ymd(date)) %>%
    summarize(patient_n = min(date[cases >= n_cases])) %>% 
    arrange(patient_n)
  head(df_patient_n)
  tail(df_patient_n)
  # add patient zero information to coronavirus data frame
  df %<>% left_join(df_patient_n, by = "Country.Region")
  df %<>% 
    group_by(Country.Region) %>%
    mutate(
    duration = interval(ymd(patient_n), ymd(date)) %>% as.numeric("days")
    ) %>% 
    ungroup()
  vars = "patient_n"
  names(vars) = paste0("patient_", n_cases)
  df %<>% rename(!!vars)
  df
}
```

Plot with aligned time series.

```{r fig.height=5, fig.width=10}
df = add_patient_n(coronavirus, n_cases = 1)
countries = c("Netherlands", "Belgium", "Germany",
              "Luxembourg", "Italy", "Switzerland", 
              "France", "Korea, South")
df_countries = df %>% filter(
  Country.Region %in% countries & type == "confirmed" & duration >= 1
  )
exclude_states = c("French Polynesia", "Saint Barthelemy", 
                   "St Martin", "Fench Guiana")
df_countries %<>% filter(!Province.State %in% exclude_states)
ggplot(df_countries, 
       aes(duration, cases, color = Country.Region)) + 
  geom_point() +
  geom_line() + 
  scale_x_continuous(breaks = seq(1, nrow(df_countries), 2)) +
  xlab("days since patient zero") +
  scale_color_few()
```

## Daily Patient Cases

Align all countries to the day when they first reached 50 confirmed cases for one single day.

```{r warning=FALSE}
n_cases = 50
df = add_patient_n(coronavirus, n_cases = n_cases)
df_countries = df %>% filter(
  Country.Region %in% countries & type == "confirmed" & duration >= 0
  )
df_countries %<>% filter(!Province.State %in% exclude_states)
```

Plot with aligned time series.

```{r}
ggplot(df_countries, 
       aes(duration, cases, color = Country.Region)) + 
  geom_point() +
  geom_line() + 
  scale_x_continuous(breaks = seq(1, nrow(df_countries), 2)) +
  xlab(paste("days since >=", n_cases, "cases on single day")) +
  scale_color_few()
```

# Session Info {.unnumbered}

```{r}
sessionInfo()
```
